---
title: "Assignment B1"
author: "Zehao Zhu"
output: github_document
Roxygen: list(markdown = TRUE)
---

### Create Function (Summary Statistics of Continuous Variable by Categorical Variable with Data Type Checking)
Let's first import the libraries.
```{r}
suppressPackageStartupMessages(library(palmerpenguins))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(testthat))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(digest))
```
When I was doing my MDA, many duplicate codes were written.\
1. In data analysis, I often have to calculate several *_summary statistics of a continuous variable grouped by a categorical variable_*.\
2. When interpreting the results, the summary statistics by non-categorical data are difficult to show and visualize as there may be too many groups. Thus, I also have to check manually if the group column is categorical and if the value column that the calculation happens on is continuous/numeric. If the variable is not categorical, I have to cut it into intervals first to make categories.\
Therefore, a lot of *_duplicate codes_* and *_manual works_* occur when calculating the summary statistics of a continuous variable grouped by a categorical column without function. Here I will create a robust function to wrap it up.
```{r}
#' @title Summarize the statistics of a numeric variable across a category
#' @details Calculate the number, mean, max, min, standard deviation, median of a numeric variable "value" across a categorical variable "group". It also checks if the group is categorical and if the value is numeric to avoid manually checking by users.
#' @param "group" should be a categorical column and "value" should be a numeric column; Naming the categorical column "group" is because it works as the categorical data to group by. The column that we perform summary statistics on is called "value" because it is numeric and is the columns that we are working on.
#' @return the summarized six statistics of value across the group
cal_stats_numeric_by_category <- function(data, group, value){
  group_numeric <- data %>%  ## check if group is numeric
    pull({{group}}) %>% 
    is.numeric
  value_numeric <- data %>%  ## check if value is numeric
    pull({{value}}) %>% 
    is.numeric
  if(group_numeric){
    stop("group is not categorical") ## stop if group is numeric
  }
  if(!value_numeric){
    stop("value is not numeric") ## stop if value is not numeric
  }
  if(data %>% nrow == 0){
    stop("empty dataset")
  }
  summarised_data <- data %>%
    group_by({{group}}) %>%
    summarise(n = n(), mean = mean({{value}}, na.rm = TRUE), max = max({{value}}, na.rm = TRUE), min = min({{value}}, na.rm = TRUE), std = sd({{value}}, na.rm = TRUE), median = median({{value}}, na.rm = TRUE))
  return(summarised_data)
}
```
### Examples to show the function
In this section, there are some examples of how it works.
```{r}
penguins %>% cal_stats_numeric_by_category(species, flipper_length_mm)
```
The above code calculates the number, mean, max, min, standard deviation, median of flipper_length groupped by species.
```{r}
gapminder %>% cal_stats_numeric_by_category(continent, lifeExp)
```
The above shows how to use this function to summarize the summary statistics on each continent.

Please note that the group variable cannot be numeric. You have to transform it into non-numeric data type before calling the function.
```{r}
new_gapminder <- gapminder
new_gapminder$year <- as.factor(new_gapminder$year)
new_gapminder %>% glimpse
```
If using the int data as group, the function will stop because it is not categorical data. After transforming the data type, it runs successfully.
```{r, error=TRUE}
gapminder %>% cal_stats_numeric_by_category(year, lifeExp) # This will run into error because the year is integer
new_gapminder %>% cal_stats_numeric_by_category(year, lifeExp)
```
If the value variable is not numeric, there is an error as well. Here is an example of error code.
```{r, error=TRUE}
gapminder %>% cal_stats_numeric_by_category(continent, country)
```
### Testing the function
First, we generate some testing data:
```{r}
test_data <- data.frame(value = 1:10, date = c("day 1", "day 1","day 1","day 1","day 1","day 2","day 2","day 2","day 2","day 2"))
test_data <- as_tibble(test_data) # The first five rows are of the same day and have values from 1 to 5. The rest five rows are of the same day and have values from 6 to 10.
n_test <- c(5, 5)
mean_test <- c((1+2+3+4+5)/5, (6+7+8+9+10)/5)
median_test <- c(3, 8)
max_test <- c(5, 10)
min_test <- c(1, 6)
std_test <- c(sd(test_data[1:5, ]$value), sd(test_data[6:10,]$value))
```

#### 1. Vector with no NA
Then, test the function on the test data:
```{r}
test_that("Testing clean data", {
  expect_length((test_data %>% cal_stats_numeric_by_category(date, value)), 7)
  expect_true(test_data %>% cal_stats_numeric_by_category(date, value)%>% pull(mean) %>% is.numeric)
  expect_equal((test_data %>% cal_stats_numeric_by_category(date, value))$"n", n_test)
  expect_equal((test_data %>% cal_stats_numeric_by_category(date, value))$"mean", mean_test)
  expect_equal((test_data %>% cal_stats_numeric_by_category(date, value))$"median", median_test)
  expect_equal((test_data %>% cal_stats_numeric_by_category(date, value))$"max", max_test)
  expect_equal((test_data %>% cal_stats_numeric_by_category(date, value))$"min", min_test)
  expect_equal((test_data %>% cal_stats_numeric_by_category(date, value))$"std", std_test)
})
```
Let's add NA to the test data:
```{r}
na_test_data <- test_data %>% add_row(value = NA, date = "day 2")
```

#### 2. Dataset that has NAs or input has NAs
```{r}
na_test_data %>% cal_stats_numeric_by_category(date, value)
test_that("Testing NA", {
  expect_error(na_test_data %>% cal_stats_numeric_by_category(NA, value))
  expect_equal((na_test_data %>% cal_stats_numeric_by_category(date, value))$"mean", mean_test)
  expect_equal((na_test_data %>% cal_stats_numeric_by_category(date, value))$"median", median_test)
  expect_equal((na_test_data %>% cal_stats_numeric_by_category(date, value))$"max", max_test)
  expect_equal((na_test_data %>% cal_stats_numeric_by_category(date, value))$"min", min_test)
  expect_equal((na_test_data %>% cal_stats_numeric_by_category(date, value))$"std", std_test)
})
```

It works fine when the data has NA.

#### 3. Variables of different types
Now, let's add more columns of different data types:
```{r}
df_test_data <- test_data %>% mutate(cur_date = Sys.Date(), float_value = value*1.0, fct_date = as.factor(date))
new_std <- sd(df_test_data$"float_value")
print(df_test_data)
```
```{r}
test_that("Testing different type", {
  expect_no_error(df_test_data %>% cal_stats_numeric_by_category(fct_date, value))
  expect_equal((df_test_data %>% cal_stats_numeric_by_category(cur_date, value))$"n", 10)
  expect_equal((df_test_data %>% cal_stats_numeric_by_category(cur_date, value))$"mean", 5.5)
  expect_equal((df_test_data %>% cal_stats_numeric_by_category(cur_date, value))$"median", 5.5)
  expect_equal((df_test_data %>% cal_stats_numeric_by_category(cur_date, value))$"max", 10)
  expect_equal((df_test_data %>% cal_stats_numeric_by_category(cur_date, value))$"min", 1)
  expect_equal((df_test_data %>% cal_stats_numeric_by_category(cur_date, value))$"std", new_std)
})
test_that("Testing invalid data type", {
  expect_error((df_test_data %>% cal_stats_numeric_by_category(value, float_value)), "group is not categorical")
  expect_error((df_test_data %>% cal_stats_numeric_by_category(date, cur_date)), "value is not numeric")
  expect_error(cal_stats_numeric_by_category(value, float_value))
})
```
#### 4. Vector of empty dataset
```{r}
empty_data <- test_data %>% filter(value == 0)
test_that("Testing empty dataset", {
  expect_error(empty_data %>% cal_stats_numeric_by_category(date, value))
})
```











